---
title: "Regression Analysis Course Project"
author: "Vicki Cocco"
date: "June 24, 2017"
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

```

## R Markdown

---
## Synopsis: 

> Part 2: Basic Inferential Data Analysis Instructionsless 
> 


### Requirement 2. Provide some a basic summary of the data.

#### Organizing the required R libraries
```{r, echo = TRUE, results='hide', message=FALSE, warning=FALSE}

        install.packages('dtplyr', repos="http://cran.rstudio.com/")
        library(dtplyr)
        install.packages('dplyr', repos="http://cran.rstudio.com/")
        library(dplyr)
        install.packages('ggplot2', repos="http://cran.rstudio.com/")
        library(ggplot2)        

```

### Data Exploratory and Modifications 

#### Example of the Raw Data after forming it to a table data structure

```{r, echo =TRUE, message=FALSE, cache=TRUE}

        # using the mtcars dataset from R
        # extracting the data used for analysis and organizing factor variables
        mpg <- mtcars$mpg
        transtype <- mtcars$am
        gears <-(mtcars$gear)
        dtCars <- cbind(mpg, transtype)
        dtCars <- cbind(dtCars, gears)
        dtCars <- tbl_df(dtCars)
        dtCars$gears <- as.factor(dtCars$gears)
        
        dtCars <- dtCars %>%
                select(mpg, gears) %>%
                mutate(Transmission = ifelse(transtype == 0, "Automatic", "Manual"))
        
        dtCars$Transmission <- as.factor(dtCars$Transmission)
        
        dtManCars <- filter(dtCars, Transmission == "Manual")
        dtAutoCars <- filter(dtCars, Transmission == "Automatic")
        
        rm(mpg, gears, transtype)
        
```




```{r, echo = TRUE, message=FALSE}
        
        knitr::kable(head(dtCars))
        
        knitr::kable(head(dtManCars))
        
        knitr::kable(head(dtAutoCars))

```

#### Showing the first regression analysis - MPG against both types of transmissions
#### All cars.


```{r, echo = TRUE, message=FALSE}

        fitAll <- lm(mpg ~ Transmission, dtCars)

        print(summary(fitAll))
        
```

#### Showing the first plot - MPG against both types of transmissions
#### All cars.

```{r, echo = TRUE, message=TRUE, fig.width=10, fig.height=8}

        fitAll <- ggplot(dtCars, aes(x = Transmission,  
                                        y = mpg,
                                        fill = Transmission))
        fitAll <- fitAll + geom_point(size = 4,
                                        shape=21,
                                        colour = "black")
        fitAll <- fitAll + geom_abline(intercept = coef(fitAll)[1],
                                        slope = coef(fitAll)[2],
                                        size = 1,
                                        color = "darkblue")
        fitAll <- fitAll + labs(x = "Automatic / Manual Transmission" ,
                                y = "Miles Per Gallon (US)", 
                                title = "Transmission Type vs MPG")
        fitAll <- fitAll + theme(plot.title=element_text(size=14, 
                                                       hjust=0.5, 
                                                       face="bold", 
                                                       colour="darkorchid4", 
                                                       vjust=-1),                                                        panel.grid.major = element_line(colour = "grey39"), 
                                panel.grid.minor = element_line(colour = "grey24"),
                                panel.background = element_blank(),
                                legend.position = "none")
        plot(fitAll)

```

#### Showing the second regression analysis - MPG against gears manual transmissions


```{r, echo = TRUE, message=FALSE}

        fitManGears <- lm(mpg ~ gears, dtAutoCars)
       
        print(summary(fitManGears))
```

#### Showing the second plot - MPG against gears of manual transmissions

```{r, echo = TRUE, message=FALSE, fig.width=10, fig.height=8}

        fitManGears <- ggplot(dtManCars, aes(x = gears, 
                                        y = mpg,
                                        fill = gears))
        fitManGears <- fitManGears + geom_point(size = 3,
                                    shape=21,
                                    colour = "black")
        fitManGears <- fitAutoGears + geom_abline(intercept = coef(fitManGears)[1],
                                                   slope = coef(fitManGears)[2],
                                                   size = 1,
                                                   color = "darkblue")
        fitManGears <- fitManGears + labs(x = "Number of Forward Gears" ,
                              y = "Miles Per Gallon (US)", 
                              title = "Manual Transmission # Forward Gears vs MPG")
        fitManGears <- fitManGears + theme(plot.title=element_text(size=14, 
                                                       hjust=0.5, 
                                                       face="bold", 
                                                       colour="darkorchid4", 
                                                       vjust=-1),                                                          panel.grid.major = element_line(colour = "grey39"), 
                               panel.grid.minor = element_line(colour = "grey24"),
                               panel.background = element_blank(),
                               legend.position = "none")
        plot(fitManGears)


```

#### Showing the third regression analysis - MPG against gears automatic transmissions

```{r, echo = TRUE, message=FALSE}

        fitAutoGears <- lm(mpg ~ gears, dtAutoCars)
       
        print(summary(fitAutoGears))
```

#### Showing the third plot - MPG against gears of automatic transmissions

```{r, echo = TRUE, message=FALSE, fig.width=10, fig.height=8}


        fitAutogears <- ggplot(dtAutoCars, aes(x = gears, 
                                    y = mpg,
                                    fill = gears))
        fitAutogears <- fitAutogears + geom_point(size = 3,
                                    shape=21,
                                    colour = "black")
        fitAutogears <- fitAutogears + geom_abline(intercept = coef(fitAutogears)[1],
                                slope = coef(fitAutogears)[2],
                                size = 1,
                                color = "darkblue")
        fitAutogears <- fitAutogears + labs(x = "Number of Forward Gears" ,
                                y = "Miles Per Gallon (US)", 
                                title = "Automatic Transmission # Forward Gears vs MPG")
        fitAutogears <- fitAutogears + theme(plot.title=element_text(size=14, 
                                                       hjust=0.5, 
                                                       face="bold", 
                                                       colour="darkorchid4", 
                                                       vjust=-1),                                                       panel.grid.major = element_line(colour = "grey39"), 
                               panel.grid.minor = element_line(colour = "grey24"),
                               panel.background = element_blank(),
                               legend.position = "none")
        
        plot(fitAutogears)

```

#### Showing the fourth regression analysis - MPG against gears automatic transmissions

```{r, echo = TRUE, message=FALSE}

        fitAllGears <- lm(mpg ~ gears, dtCars)
       
        print(summary(fitAllGears))
```

#### Showing the third plot - MPG against gears of automatic transmissions

```{r, echo = FALSE, message=FALSE, fig.width=10, fig.height=8}

        fitAllGears <- ggplot(dtCars, aes(x = gears, 
                                                y = mpg,
                                                fill = Transmission))
        fitAllGears <- fitAllGears + geom_point(size = 3,
                                                shape=21,
                                                colour = "black")
        fitAllGears <- fitAllGears + geom_abline(intercept = coef(fitAllGears)[1],
                                                   slope = coef(fitAllGears)[2],
                                                   size = 1,
                                                   color = "darkblue")
        fitAllGears <- fitAllGears + labs(x = "Number of Forward Gears" ,
                                        y = "Miles Per Gallon (US)", 
                                        title = "Number of Forward Gears vs MPG")
        fitAllGears <- fitAllGears + theme(plot.title=element_text(size=14, 
                                                       hjust=0.5, 
                                                       face="bold", 
                                                       colour="darkorchid4", 
                                                       vjust=-1),                                                       panel.grid.major = element_line(colour = "grey39"), 
                             panel.grid.minor = element_line(colour = "grey24"),
                             panel.background = element_blank())
                            
        plot(fitAllGears)
```

### Fourth Hypothesis
#### H_0 - Status Quo Toothgrowth Length
#### H_a - A Dose of 1, through either delivery method, improves tooth growth
#### T-Test Length ~ Dose @ 2
#### Outcome : The p-value is small showing that it is unlikely these results would be
#### obtained without the impact of the dose.
#### The alternative hypothesis should be accepted.

```{r, echo = FALSE, message=FALSE}
        #tblNull2 <- filter(tblNull, Dose != 2)
        #tblDose2 <- filter(tblDose, Dose == 2)
        #print(t.test(tblNull2$Length, tblDose2$Length))

```
### Conclusions:
### The delivery method of Vitamin C has no impact on tooth growth.
### The dose level of Vitamin C has varying impacts, with the dose level of 
### 2.0 mg/day having the greatest improvement on tooth growth.
